---
title: "Simulations In Statistics"
author: "Jerome Cordjotse"
date: "5/18/2020"
output: pdf_document
---

```{r setup, include=FALSE}
library(Hmisc)
knitr::opts_chunk$set(echo = TRUE, fig.width = 8)
set.seed(2020)
```

## Overview {#intro}

The mean and standard deviation for an exponential distribution with lambda, $\lambda$, 0.2 is 5 since mean = $\frac{1}{\lambda}$ and standard deviation = $\sqrt(var) = \sqrt(\frac{1}{\lambda^{2}}) = \frac{1}{\lambda}$. In this experiment, the Central Limiting Theory, CTL would be used to figure out this mean through experiments and simulations. Detailed code and simulation explanations can be found here in the [link to my repo](https://github.com/jecordjotse/StatsInfAssignment).



## Simulations {#sima}
Here a thousand simulations where made, each retrieving the mean of 40 exponentials. This distribution is stored in a $42 \times 1000$ matrix with columns storing the sample of size 40 and the last rows, 41 & 42, sotres the mean of the samples from row 1 to 40 and their respective variance respectively. For a thousand simulations the column number becomes 1000.
```{r }
lambda = 0.2
mean <- 1/lambda           # Theoritical Mean
sd <- 1/lambda             # Theoritical Standard Deviation
var <- sd^2                # and variance
## To make a 1000 simulations with a for loop
sim_mat <- NULL
for (i in 1 : 1000){
  expObs <- rexp(40,0.2)
  sim_mat <- cbind(         # Adding simulations by
      sim_mat,              # attaching new columns
        c(
          expObs,           # 40 exponentials
          mean(expObs),     # adding their mean to at row 41
          var(expObs)       # adding their variance at row 42
          )
        )
}

sampleMean <- sim_mat[41,]
sample_Variance_Distribution <- sim_mat[42,]
summary(sampleMean)
```

## Sample Mean vs Theoretical Mean

Knowing that the sample mean is a good estimator of the population mean. So ploting the histogram of means shows this as it balances around 5.

```{r  echo=FALSE}
hist(sampleMean, main="Histogram of Sample Mean", xlab="Sample Mean")
abline(v=mean(sampleMean), col="blue", lwd = '2')
```

This can be confirmed as the average of sample mean is below which is close to 5.

```{r }
mean(sampleMean)
```


## Sample Variance vs Theoretical Variance
As seen with the distribution of sample means, the sample variance estimates the population variance. Taking average yields

```{r }
mean(sample_Variance_Distribution)
```

```{r echo=FALSE}
d <- density(sample_Variance_Distribution)
plot(d, xlab="Sample Variance", main="Desnsity Distribution of Sample Variance")
axis(1,at=round(mean(sample_Variance_Distribution),2),las = 2)
abline(v=mean(sample_Variance_Distribution), col="blue")
```

## Distribution
``` {r echo=FALSE}
hist(sampleMean, freq = FALSE, main = "Histogram of Distribution of 1000 averages") # histograph of the distribution of 1000 sample means
curve(dnorm(x,mean=mean(sampleMean),sd=sd(sampleMean)), add=TRUE, col="red") # The Distribution is aproximately a normal
abline(v=mean(sampleMean), col="blue", lwd = '2')
```
The CLT states that the distribution of averages of iid variables becomes that of a standard
normal as the sample size increases. It can be seen than the Figure above shows an approximation to a normal as the bell-curve shape overlays it.

